---
title: "Yemen Cholera Data Exploration"
author: "Andrea Pedot"
date: "2024-08-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preliminary steps

The project is written according to the LAIMS analysis template, thus most of the objects can be retrieved through Quarto.

```{r}
db <- tar_read(ycc)

dbpop <- db |>
  ## generate population counts
  dplyr::group_by(pcode) |>
  dplyr::summarise(pop = median( cases/(attack_abs/1000)) )
  
a <- db |>
  dplyr::left_join(dbpop, by = dplyr::join_by(pcode)) |>
  dplyr::group_by(pcode) |>
  dplyr::mutate(
    dplyr::across(
      cases:deaths,
      ~ .x - dplyr::lag(.x, default = 4L),
      .names = "{.col}_4w"
    )) |>
  dplyr::mutate(cfr_4w = deaths_4w / cases_4w,
                attack_4w = cases_4w / pop
                ) |>
  dplyr::ungroup()
  



```


Also, relevant libraries are loaded

```{r}
library(tidyverse)
ycc <- tar_read(ycc)
yem_shp <- tar_read(yem_shp)
yemen <- tar_read(yemen)

yemen |>
  ggplot() +
    geom_sf(aes(fill = ADM1_EN))
```

Data should be grouped by the Epi-Week where they have been recorded.

## General description of the Epidemics

The dataset contains the description of Cholera cases in Yemen from 22nd May 2017 to 18th February 2018, with no missing data.

```{r, echo = FALSE}
ycc |>
  ggplot() +
  geom_line(aes(x = date, y = cases, color = govt)) +
  facet_wrap(~govt)
```

Focusing on a single district, the new cases can be visualised.

```{r, echo = FALSE}
ycc |>
  filter(govt == "Dhamar") |>
  mutate(new_cases = cases - lag(cases)) |>
  ggplot() +
  geom_line(aes(x = date, y = new_cases))
```

Then, the Yemen shapefile can be imported.

```{r}
yemshp <- tar_read(yem_shp)
```
