---
title: "Yemen Cholera Data Exploration"
author: "Andrea Pedot"
date: "2024-08-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preliminary steps

The project is written according to the LAIMS analysis template, thus most of the objects can be retrieved through Quarto. Also, relevant libraries are loaded.

```{r}
tar_load(ycc_lag)
tar_load(yemen_4w)

library(tidyverse)
library(sf)
```




```{r}
ycc_lag |>
  dplyr::group_by(epi_date) |>
  dplyr::mutate(
    # Expected counts
    cases_exp = cases * ( sum(cases)/ sum(pop)),
    deaths_exp = cases * ( sum(deaths)/ sum(pop)),
    cfr_exp = deaths_exp / cases_exp,
    attack_exp = cases_exp / pop,
    # Target variable
    target_status = lead(status)
  ) |>
  dplyr::filter(epi_date == "2017-51") 

```

```{r}
d <- tar_read(yemen_4w) |>
  dplyr::select("status-2018-1","cases_exp-2017-51",
                "attack_exp-2017-51","status-2017-51") |>
    dplyr::rename(
      "target_status[]" = "status-2018-1",
      "cases_exp[]" = "cases_exp-2017-51",
      "attack_exp[]" = "attack_exp-2017-51",
      "status[]" = "status-2017-51")
d <- format(d$`attack_exp[]`, scientific = FALSE)
```

Data should be grouped by the Epi-Week where they have been recorded.

## General description of the Epidemics

The dataset contains the description of Cholera cases in Yemen from 22nd May 2017 to 18th February 2018, with no missing data.

```{r, echo = FALSE}
ycc |>
  ggplot() +
  geom_line(aes(x = date, y = cases, color = govt)) +
  facet_wrap(~govt)
```

Focusing on a single district, the new cases can be visualised.

```{r, echo = FALSE}
ycc |>
  filter(govt == "Dhamar") |>
  mutate(new_cases = cases - lag(cases)) |>
  ggplot() +
  geom_line(aes(x = date, y = new_cases))
```

Then, the Yemen shapefile can be imported.

```{r}
yemshp <- tar_read(yem_shp)
```
