---
title: "Yemen Cholera Data Exploration"
author: "Andrea Pedot"
date: "2024-08-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preliminary steps

The project is written according to the LAIMS analysis template, thus most of the objects can be retrieved through Quarto. Also, relevant libraries are loaded.

```{r}
tar_load(ycc)
tar_load(ycc_lag)
tar_load(yemen)
tar_load(yemen_4w)

library(tidyverse)
library(sf)
```




```{r}
status_color <- function(status) {
  dplyr::case_match(status,
    0 ~ "green",
    1 ~ "orange",
    2 ~ "red"
  )
}

yemen_4w |>
  ggplot2::ggplot() +
  ggplot2::geom_sf(aes(fill = `status-2017-40`)) +
  ggplot2::labs(title = "Cholera region Status") +
  ggplot2::theme_light()

```

Data should be grouped by the Epi-Week where they have been recorded.

## General description of the Epidemics

The dataset contains the description of Cholera cases in Yemen from 22nd May 2017 to 18th February 2018, with no missing data.

```{r, echo = FALSE}
ycc |>
  ggplot() +
  geom_line(aes(x = date, y = cases, color = govt)) +
  facet_wrap(~govt)
```

Focusing on a single district, the new cases can be visualised.

```{r, echo = FALSE}
ycc |>
  filter(govt == "Dhamar") |>
  mutate(new_cases = cases - lag(cases)) |>
  ggplot() +
  geom_line(aes(x = date, y = new_cases))
```

Then, the Yemen shapefile can be imported.

```{r}
yemshp <- tar_read(yem_shp)
```
