---
title: "Yemen Cholera Data Exploration"
author: "Andrea Pedot"
date: "2024-08-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preliminary steps

The project is written according to the LAIMS analysis template, thus most of the objects can be retrieved through Quarto.

Also, relevant libraries are loaded

```{r}
library(tidyverse)
ycc <- tar_read(ycc)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## General description of the Epidemics

The dataset contains the description of Cholera cases in Yemen from 22nd May 2017 to 18th February 2018, with no missing data.

```{r}
ycc |>
  ggplot() +
  geom_line(aes(x = date, y = cases)) +
  facet_wrap(~govt)
```

Focusing on a single district, the new cases can be visualised.

```{r}
ycc |>
  mutate(new_cases = lag(cases) - cases, .by = govt) |>
  filter(govt == "Dhamar") |>
  ggplot() +
  geom_line(aes(x = date, y = new_cases))
```

